Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-12 paper.tex
wordings

In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
benchmarked against several competitors. In this paper, we instead focus on an
in-depth discussion and analysis of the design choices used in building the
library. First, in section \ref{sec:core-api}, we present the central application
programming interface (API) and then describe, in section \ref{sec:advanced-api},
advanced API mechanisms built on this core API. Section
\ref{sec:implementation} describes the implementation trade-offs and analyzes
the limitations of the technologies we rely on while section
\ref{sec:comparison} then compares \emph{scikit-learn} to other major projects in
terms of API. Finally, conclusions and directions of future works are included
in section \ref{sec:conclusions}.

@@

\begin{description}
  \item[Inspection.]
       Constructor parameters
       and parameter values determined by learning algorithms
       are stored as public attributes. All of them are documented in a consistent manner.
       % Gilles: Do we really need to bother the reader with such details?
       % (The Python language does not actually distinguish
       % between public and private attributes of objects,
       % but by convention, any identifier
       % prefixed with an underscore (\texttt{\_})
       % is considered ``a non-public part of the API''; \citealp{pythontut}.)
  \item[Non-proliferation of classes.]
       Learning algorithms are objects of custom classes,
       but datasets and hyperparameter names and values
       are represented as standard Python, NumPy or SciPy classes
       rather than new ones whenever feasible.
       % Gilles: I don't understand the phrase below
       % Custom classes adhere to Python conventions and protocols
       % for constructs such as iteration over the folds
       % in a $k$-fold cross validation.
       This should keep \sklearn easy to use
       and easy to combine with other libraries.
       % XXX do we need to elaborate more on this, or do we expect everyone
       % to be an experienced Java/adapter pattern hater?
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parameterized on other algorithms.
       To be able to implement and use these algorithms in a simple way,
       all learners (basic or composed) are usable with a consistent set of methods.
       % Gilles: let elaborate later on the syntax
       % and a simple syntax is implemented
       % for setting and retrieving parameters of sub-estimators.
       % XXX need to describe this!
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

>>>>

In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
benchmarked against several competitors. In this paper, we instead focus on an
in-depth discussion and analysis of the design choices used in building the
library. First, in section \ref{sec:core-api}, we present the central application
programming interface (API) and then describe, in section \ref{sec:advanced-api},
advanced API mechanisms built on this core API. Section
\ref{sec:implementation} describes implementation trade-offs and analyzes
limitations of the technologies we rely on, while section
\ref{sec:comparison} then compares \emph{scikit-learn} to other major projects in
terms of API. Finally, conclusions and directions of future works are included
in section \ref{sec:conclusions}.

@@

\begin{description}
  \item[Consistency.]
       All objects (basic or composed) share a consistent interface composed of
       a limited set of methods.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored as public attributes. All of them are documented in
       a consistent manner.
       % Gilles: Do we really need to bother the reader with such details?
       % (The Python language does not actually distinguish
       % between public and private attributes of objects,
       % but by convention, any identifier
       % prefixed with an underscore (\texttt{\_})
       % is considered ``a non-public part of the API''; \citealp{pythontut}.)
  \item[Non-proliferation of classes.]
       Only learning algorithms are objects of custom classes,
       Datasets and hyperparameter names and values
       are represented as standard Python, NumPy or SciPy classes
       rather than new ones whenever feasible.
       % Gilles: I don't understand the phrase below
       % Custom classes adhere to Python conventions and protocols
       % for constructs such as iteration over the folds
       % in a $k$-fold cross validation.
       This should keep \sklearn easy to use
       and easy to combine with other libraries.
       % XXX do we need to elaborate more on this, or do we expect everyone
       % to be an experienced Java/adapter pattern hater?
  % Gilles: this is actually not a design principle
  % \item[Composition.]
  %      Many machine learning tasks are expressible
  %      as sequences or combinations of transformations to data.
  %      Some learning algorithms are also naturally viewed
  %      as meta-algorithms parameterized on other algorithms.
  %      To be able to implement and use these algorithms in a simple way,
  %      % Gilles: let elaborate later on the syntax
  %      % and a simple syntax is implemented
  %      % for setting and retrieving parameters of sub-estimators.
  %      % XXX need to describe this!
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}
